---
title: "Peer-graded Assignment: Getting and Cleaning Data Course Project"
author: "Chuck Chekuri"
date: "2/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Codebook

This codebook explains the code in run_analysis.R

Files were downloaded using the following R code.
The run_analysis.R script assumes the working directory to 
be the "UCI HAR Dataset" directory.

> setwd("~/GitHub/datasciencecoursera/DataGatheringAssignment")
> zipurl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
> download.file(zipurl, destfile = "data.zip", method="curl")
> unzip("data.zip")    
> setwd("./UCI HAR Dataset")


# loading libraries
> library(data.table)
> library(dplyr)

## read only what is necessary
Features has all the 561 columns but we need only 6
Read only the first 6 of 561 columns as needed by the assignment
 3 means and 3 std deviations and set the rest to NULL
cols2Read creates an list to only read those columsn

> cols2Read <- c(rep(NA,6),rep("NULL",555))

Similarly read only the feature names we need ( first 6 )
> features <- read.table('./features.txt', header=FALSE, 
>		       stringsAsFactors = FALSE, nrows = 6)[,2]

### Next we read in the activity labels
> act_lab <-read.table("./activity_labels.txt")

### read test data 
> subj_test <- read.table('./test/subject_test.txt')
> y_test <- read.table('./test/y_test.txt')
> x_test <- read.table('./test/X_test.txt', colClasses = cols2Read)

### read train data
> subj_train <- read.table('./train/subject_train.txt')
> y_train <- read.table('./train/y_train.txt')
> x_train <- read.table('./train/X_train.txt', colClasses = cols2Read)

### convert these to data tables for easy merging
> act_lab <- as.data.table(act_lab)
> y_test <-as.data.table(y_test)
> y_train <-as.data.table(y_train)

### bind all columns and merge activity labels
> test_recs <- cbind(subj_test, merge(y_test, act_lab)[,2], x_test)
> train_recs <- cbind(subj_train, merge(y_train, act_lab)[,2], x_train)

### bind all rows 
> allrecs <- as.data.table(rbind(train_recs, test_recs))

### label columns
> colnames(allrecs) <- c("subject", "activity", features)
> allrecs$subject <- as.factor(allrecs$subject)

### write all the rows to measurements.csv
> write.table(allrecs, "measurements.csv", row.names=FALSE, quote=FALSE,sep=",")

### create second tidy data of grouped means
> grouped_means <- allrecs %>% 
>   group_by(activity, subject) %>% 
>   summarize( 
>   	AverageBodyAccMeanX=mean(`tBodyAcc-mean()-X`),
>   	AverageBodyAccMeanY=mean(`tBodyAcc-mean()-Y`),
>   	AverageBodyAccMeanZ=mean(`tBodyAcc-mean()-Z`),
>  	AverageBodyAccStdX=mean(`tBodyAcc-std()-X`),
>   	AverageBodyAccStdY=mean(`tBodyAcc-std()-Y`),
>   	AverageBodyAccStdZ=mean(`tBodyAcc-std()-Z`)
>   	)

### write grouped means to grouped_means.csv
> write.table(grouped_means, "grouped_means.csv", row.names=FALSE, quote=FALSE,sep=",")

## Output
### Measurement.csv
```{r output, include=TRUE}
x <- read.table("measurements.csv", nrow=10)
head(x)
```
### grouped_means.csv
```{r groupedMeans, include=TRUE}
x <- read.table("grouped_means.csv", nrow=10)
head(x)
```
